FictionEro2 - Affective Impact of Fiction

### 2025-03-18 study2/analysis/1_cleaning.qmd
minor
analysis scripts updated

df |>
  select(Feedback_AllRealConfidence, Feedback_AllFakeConfidence) |>
  pivot_longer(everything(), names_to = "Authenticity", values_to = "Confidence") |>
  drop_na() |>
  mutate(Authenticity = str_remove(Authenticity, "Feedback_"),
         Authenticity = str_replace(Authenticity, "AllRealConfidence", "All images were real"),
         Authenticity = str_replace(Authenticity, "AllFakeConfidence", "All images were fake")) |>
  ggplot(aes(x = Confidence, fill= Authenticity)) +
  geom_bar() +
  labs(x="Confidence (0-5)", y = "Number of Participants", title = "Confidence on the authenticity of the images") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.title.x = element_blank())
```

@@

data.frame(Comments = df$Feedback_Text) |>
  filter(!tolower(Comments) %in% c(NA, "no", "no comments", "nothing", "none", "", "no.")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE)

@@

```{r}
#| echo: true
#| code-fold: false

@@

@@

`r length(outliers$arousal)` participant had a negative correlation between arousal and enticement, and had a negative correlation between arousal and valance.

@@

```{r}
p_edu <- df |>
  mutate(Education = fct_relevel(Education, "Other", "High School", "Bachelor", "Master", "Doctorate")) |> 
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1), angle = 45, hjust = 1),
    axis.title.x = element_blank()
  )
p_edu
```

@@

>>>>

df |>
  select(Feedback_AllRealConfidence, Feedback_AllFakeConfidence) |>
  pivot_longer(everything(), names_to = "Authenticity", values_to = "Confidence") |>
  drop_na() |>
  mutate(Authenticity = str_remove(Authenticity, "Feedback_"),
         Authenticity = str_replace(Authenticity, "AllRealConfidence", "All images were real"),
         Authenticity = str_replace(Authenticity, "AllFakeConfidence", "All images were fake")) |>
  ggplot(aes(x = Confidence, fill= Authenticity)) +
  geom_bar() +
  scale_x_continuous(n.breaks = 6) +
  labs(x="Confidence (0-5)", y = "Number of Participants", title = "Confidence on the authenticity of the images") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.title.x = element_blank())
```

@@

data.frame(Comments = df$Feedback_Text) |>
  filter(!tolower(Comments) %in% c(NA, "no", "no comments", "nothing", "none", "", "no.", "no. ", "n/a", "n/a - thanks", "thanks")) |> 
  gt::gt() |> 
  gt::opt_stylize() |> 
  gt::opt_interactive(use_compact_mode = TRUE)

@@

```{r}
#| echo: true
#| code-fold: true

@@

### BAIT Questionnaire Duration
#
```{r}
#| fig-width: 15
#| fig-height: 15

df |>
  mutate(Participant = fct_reorder(Participant, Bait_Duration),
         Category = ifelse(Bait_Duration > 5, "extra", "ok"),
         Duration = ifelse(Bait_Duration > 5, 5, Bait_Duration),
         Group = ifelse(Participant %in% outliers, "Outlier", "ok")) |>
  ggplot(aes(y = Participant, x = Duration)) +
  geom_point(aes(color = Group, shape = Category)) +
  geom_vline(xintercept = median(df$Bait_Duration), color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0.5, color = "orange", linetype = "dotted") +
  scale_shape_manual(values = c("extra" = 3, ok = 19)) +
  scale_color_manual(values = c("Outlier" = "red", ok = "black"), guide="none") +
  guides(color = "none", shape = "none") +
  ggside::geom_xsidedensity(fill = "#9C27B0", color=NA) +
  ggside::scale_xsidey_continuous(expand = c(0, 0)) +
  labs(
    title = "Questionnaire Completion Time",
    x = "Duration (in minutes)",
    y = "Participant"
  )  +
  theme_modern() +
  ggside::theme_ggside_void() +
  theme(ggside.panel.scale = .3,
        panel.border = element_blank(),
        axis.ticks.y = element_blank(),
          axis.text.y = element_blank())

outliers$bait_duration <- filter(df, Bait_Duration < 0.5)$Participant
df[df$Participant %in% outliers$bait_duration, str_detect(names(df), "BAIT_\\d+")] <- NA
```

`r length(outliers$bait_duration)` participants completed the questionnaire in less than 30 seconds.

### Response to Erotic Stimuli

```{r}
# Single arousal response (0)
outliers$novariability <- summarize(dftask, n = length(unique(Arousal)), .by="Participant") |> 
  filter(n == 1) |> 
  select(Participant) |> 
  pull()

df <- filter(df, !Participant %in% outliers$novariability)
dftask <- filter(dftask, !Participant %in% outliers$novariability)
```

`r length(outliers$novariability)` participants showed no variation in their arousal response (did not move the scales).

@@

`r length(outliers$arousal)` participant had a negative correlation between arousal and enticement, AND had a negative correlation between arousal and valance.

```{r}
df <- filter(df, !Participant %in% outliers$arousal)
dftask <- filter(dftask, !Participant %in% outliers$arousal)
```

@@

```{r}
p_edu <- df |>
  mutate(Education = fct_relevel(Education, "High school", "Bachelor", "Master", "Doctorate", "Other"))|> 
  ggplot(aes(x = Education)) +
  geom_bar(aes(fill = Education)) +
  scale_y_continuous(expand = c(0, 0), breaks= scales::pretty_breaks()) +
  scale_fill_viridis_d(guide = "none") +
  labs(title = "Education", y = "Number of Participants", subtitle = "Participants per achieved education level") +
  theme_modern(axis.title.space = 15) +
  theme(
    plot.title = element_text(size = rel(1.2), face = "bold", hjust = 0),
    plot.subtitle = element_text(size = rel(1.2), vjust = 7),
    axis.text.y = element_text(size = rel(1.1)),
    axis.text.x = element_text(size = rel(1.1), angle = 45, hjust = 1),
    axis.title.x = element_blank()
  )
p_edu
```

@@

## Save

```{r}
#| code-fold: false

# setdiff(unique(df$Participant), unique(dftask$Participant))
write.csv(df, "../data/data_participants.csv", row.names = FALSE)
write.csv(dftask, "../data/data.csv", row.names = FALSE)
```

### 2025-03-18 study2/analysis/2_emotion.qmd
analysis scripts updated

```{r}
dfsub <- read.csv("../data/data_participants.csv")
df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Mobile, starts_with(c("Feedback_","BAIT")), COPS_Frequency_2, GenderualActivity,
           -Feedback_Text),
    by = "Participant"
    ) |> 
  mutate(Condition = fct_relevel(Condition, "Reality", "Fiction"),
         Gender =  fct_relevel(Gender, "Male", "Female", "Other"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency_2)),
         GenderualActivity_num = as.numeric(as.factor(GenderualActivity)),
         ConditionBelief = case_when(
           Condition == "Reality" & Realness > 0.5 ~ "True",
           Condition == "AI-Fiction" & Realness < 0.5 ~ "True",
           .default = "False",
         )) |> 
  mutate(Condition = case_when(
    Condition == "Fiction" ~ "AI-Generated",
    Condition == "Reality" ~ "Photograph"
  )) |>
  mutate(across(starts_with("Feedback_")& !matches("Feedback_AllRealConfidence|Feedback_AllFakeConfidence|Feedback_Enjoyment"), function(x) {fct_relevel(x, "False", "True")}))

@@

**Note to dom** beta model does not make sense anymore, what else should we do?

@@

@@

# the model with participants and stimuli as random effecs seems to be the best
```

This model evaluate ths the effects of Gender on Arousal, accounting for random variability due to participants and the stimuli (i.e., random effects).

The analysis revealed that on average **women rate erotic stimuli significantly lower in arousal than men**, while participants identifying as 'other' rate on average erotic stimuli higher than men, however this is not significant.

```{r}
m_a <- glmmTMB::glmmTMB(Arousal ~ Gender + (1|Participant) + (1|Stimulus ),
                      data=df,
                      control = glmmTMB::glmmTMBControl(parallel = 8))
                      # control = glmmTMB::glmmTMBControl(optimizer=stats::optim, # nlminb, optim
                      #                                   optArgs = list(method = "BFGS"),  #Nelder-Mead, BFGS, CG
                      #                                   parallel = 8))

@@

```{r}
valence <- df |> 
  ggplot(aes(x=Enticing)) +
  geom_histogram(fill="grey", position = "identity", binwidth = 0.5) +
  scale_y_sqrt(expand = c(0, 0)) +
  scale_x_continuous(breaks = (0:6), labels = c("Not at all", "1", "2", "3", "4", "5", "Very much")) +
  labs(title = "Valence") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"),
        legend.position = "none")

@@

```{r}
m_e <- glmmTMB::glmmTMB(Enticing ~ Gender + (1|Participant) + (1|Stimulus ),
                      data=df,
                      control = glmmTMB::glmmTMBControl(parallel = 8))
                      # control = glmmTMB::glmmTMBControl(optimizer=stats::optim, # nlminb, optim
                      #                                   optArgs = list(method = "BFGS"),  #Nelder-Mead, BFGS, CG
                      #                                   parallel = 8))

@@

@@

The analysis revealed that on average **women rate erotic stimuli significantly lower in arousal than men**, while participants identifying as 'other' rate on average erotic stimuli higher than men,. None of these differences were statistically significant.

```{r}
m_v <- glmmTMB::glmmTMB(Valence ~ Gender + (1|Participant) + (1|Stimulus ),
                      data=df,
                      control = glmmTMB::glmmTMBControl(parallel = 8))
                      # control = glmmTMB::glmmTMBControl(optimizer=stats::optim, # nlminb, optim
                      #                                   optArgs = list(method = "BFGS"),  #Nelder-Mead, BFGS, CG
                      #                                   parallel = 8))

@@

@@

@@

>>>>

```{r}
dfsub <- read.csv("../data/data_participants.csv")

df <- read.csv("../data/data.csv") |> 
  right_join(
    select(dfsub, Participant, Mobile, starts_with(c("Feedback_","BAIT")), COPS_Frequency_2, SexualActivity,
           -Feedback_Text),
    by = "Participant"
    ) |> 
  mutate(Condition = case_when(
    Condition == "Fiction" ~ "AI-Generated",
    Condition == "Reality" ~ "Photograph"
  )) |>
  mutate(Condition = fct_relevel(Condition, "Photograph", "AI-Generated"),
         Gender =  fct_relevel(Gender, "Male", "Female", "Other"),
         PornFrequency = as.numeric(as.factor(COPS_Frequency_2)),
         GenderualActivity_num = as.numeric(as.factor(SexualActivity)),
         ConditionBelief = case_when(
           Condition == "Photograph" & Realness > 0 ~ "True",
           Condition == "AI-Generated" & Realness < 0 ~ "True",
           .default = "False",
         )) |>
  mutate(across(starts_with("Feedback_")& !matches("Feedback_AllRealConfidence|Feedback_AllFakeConfidence|Feedback_Enjoyment"), function(x) {fct_relevel(x, "False", "True")}))

@@

@@

Arousal ratings were relatively well-distributed across the scale, with the highest frequency for "Not at all" and a gradual decrease across higher ratings.

The pattern of arousal ratings was similar regardless of whether participants believed the images were AI-generated or real. Among male participants, arousal ratings were relatively evenly distributed but tended to cluster around the middle of the scale, independent of their belief about the images' origin. Female participants, on the other hand, showed a pronounced peak at the lowest arousal rating, irrespective of their perception of the images as AI-generated. Participants with other gender identities provided fewer responses overall, but their arousal ratings generally leaned toward moderate levels, particularly in the "Not Believed" condition.

@@

# the model with participants only as random effecs seems to be the best
```

This model evaluate ths the effects of Gender on Arousal, accounting for random variability due to participants (i.e., random effects). The analysis revealed that on average **women rate erotic stimuli significantly lower in arousal than men**, while participants identifying as 'other' rate on average erotic stimuli higher than men, however this is not significant.

```{r}
m_a <- glmmTMB::glmmTMB(Arousal ~ Gender + (1|Participant),
                      data=df, family=glmmTMB::ordbeta(),
                      control = glmmTMB::glmmTMBControl(parallel = 8))
                      # control = glmmTMB::glmmTMBControl(optimizer=stats::optim, # nlminb, optim
                      #                                   optArgs = list(method = "BFGS"),  #Nelder-Mead, BFGS, CG
                      #                                   parallel = 8))

@@

Enticement ratings were relatively well-distributed across the scale.

Male participants predominantly provided responses in the mid-to-high enticement range across both conditions. Female participants showed a pronounced peak at the lowest enticement rating when they did not believe the image was AI-generated, with a similar but less pronounced pattern in the believed condition. Participants identifying as 'Other' responses were more concentrated around the middle of the scale.

```{r}
enticement <- df |> 
  ggplot(aes(x=Enticing)) +
  geom_histogram(fill="grey", position = "identity", binwidth = 0.5) +
  scale_y_sqrt(expand = c(0, 0)) +
  scale_x_continuous(breaks = (0:6), labels = c("Not at all", "1", "2", "3", "4", "5", "Very much")) +
  labs(title = "Enticement") +
  theme_abyss() +
  theme(plot.title = element_text(hjust = 0.5, face="bold"),
        legend.position = "none")

enticement
```

@@

```{r}
m_e <- glmmTMB::glmmTMB(Enticing ~ Gender + (1|Participant),
                      data=df,
                      control = glmmTMB::glmmTMBControl(parallel = 8))
                      # control = glmmTMB::glmmTMBControl(optimizer=stats::optim, # nlminb, optim
                      #                                   optArgs = list(method = "BFGS"),  #Nelder-Mead, BFGS, CG
                      #                                   parallel = 8))

@@

Valence ratings were concetrated more around the mid-to-high valence.

Participants showed the same pattern of results regardless of whether they believed the images to be Ai-generated or not. 
Males tended to to have the majority of their responses concentrated in the mid-to-high valence range. Female majoity valence ratings were on on the lower (0) and midle end of the scale (3/4).

@@

The analysis revealed that on average **women rate erotic stimuli significantly lower in valence than men**, while participants identifying as 'other' rate on average erotic stimuli higher in valence than men but these was not statistically significant.

```{r}
m_v <- glmmTMB::glmmTMB(Valence ~ Gender + (1|Participant),
                      data=df,
                      control = glmmTMB::glmmTMBControl(parallel = 8))
                      # control = glmmTMB::glmmTMBControl(optimizer=stats::optim, # nlminb, optim
                      #                                   optArgs = list(method = "BFGS"),  #Nelder-Mead, BFGS, CG
                      #                                   parallel = 8))

@@

In contrast to [study1](https://realitybending.github.io/FictionEro/study1/analysis/2_emotions.html) arousal ratings differ by whether the images were preceeded by 'photograph' or by 'ai-generated'.

@@

This model examines the effect of Gender and condition on Arousal, accounting for random variability due to participants and items (i.e., random effects).

<!-- The intercept is Male and photograph.  -->
Women rated 'photographs' significantly lower in arousal than men.
Men rated 'photographs' significantly higher in arousal than 'AI-generated' images.

@@

m_arousal <- glmmTMB::glmmTMB(Arousal ~ Gender/ Condition + 
                                (1| Participant) + (1|Stimulus), data=df, control = glmmTMB::glmmTMBControl(parallel = 8))

results_table(m_arousal)

# marginaleffects::avg_slopes(m_arousal, variables="Condition", by="Gender", re.form=NA) |> 
#   results_table()
